---
title: "Import raw data"
author: "Joao Marreiros"
date: "`r Sys.time()`"
output: 
  pdf_document: default
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```

___

__Brief description of the script__

This R markdown document imports and formats the output of the resulting CSV file from the computing ISO 25178-2 parameters in ConfoMap. These data is part of the manuscript: *Dubreuil et al. A ‘family of wear’: Exploring use-wear patterns on ad hoc burnishing tools*

The script includes three steps:

1. Reads in the original CSV-file   
2. Formats the data   
3. Writes an XLSX-file and save an R object ready for further analysis in R

This R project and respective scripts follow the procedures described by Marwick et al. 2017. 

The authors would like to thank Ivan Calandra and Lisa Schunk for their help and contribution on several chunks of code included here in the script (pieces of code are also adapated from Calandra et al. 2019, Pedergnana et al. 2020a, 2020b).

To compile this markdown document do not delete or move files from their original folders.

For any questions, comments and inputs, please contact:

Joao Marreiros, marreiros@rgzm.de

___


# Load libraries
```{r}
library(openxlsx)
library(tools)
library(R.utils)
library(chron)

dir_in <- "analysis/raw_data/"
dir_out <- "analysis/derived_data/"
```

___

# Get file names, path and info
```{r}
data_file <- list.files(dir_in, pattern = "\\.csv$", full.names = TRUE)
md5_in <- md5sum(data_file)
```


# Import and read the original CSV-file
```{r}
imp_data <- read.csv(data_file, header = FALSE, na.strings = "*****", encoding = "latin1")
str(imp_data)
```

___

# Format data
## Keep only interesting columns and rows
```{r}
# keeps only the columns and rows of interest for the analysis 
data_keep_col <- c(1:2, 16:17, 20:22, 24:26, 29:32, 50:83)
data_keep_rows <- which(imp_data[[1]] != "#")
data_keep <- imp_data[data_keep_rows, data_keep_col]
```

## Add headers
```{r}
head_data_keep <- unlist(imp_data[2, data_keep_col]) 
colnames(data_keep) <- gsub("\\.+", "\\.", make.names(head_data_keep))
colnames(data_keep) <- gsub("\\.$", "", colnames(data_keep)) 
```

## Identify results using frame numbers
```{r}
# combines the results from the different analysis based on the column numbers 
# (ID from MountainsMAp)
frames <- as.numeric(unlist(imp_data[1, data_keep_col]))
ID <- which(frames == 6)[-(1:2)]
ISO <- which(frames == 17)
furrow <- which(frames == 18)
diriso <- which(frames %in% 19:20)
SSFA <- which(frames %in% 21:22)
```


## Shorten the names for parameters
```{r}
# keeps only the important information of the headers 
colnames(data_keep)[ISO] <- sapply(strsplit(names(data_keep)[ISO], ".", fixed = TRUE), `[[`, 1)
colnames(data_keep)[SSFA] <- gsub("^([A-Za-z0-9]+\\.)+", "", colnames(data_keep)[SSFA])
```


## Save units
```{r}
# takes the units which were part of the headers and separates them; creates a data frame
var_num <- c(ID, ISO, furrow, diriso, SSFA)
# extracts 'unit' line for considered columns
units_var <- unlist(imp_data[3, data_keep_col])[var_num] 
# gets names associated to the units
names(units_var) <- head_data_keep[var_num] 
# puts all of it into a data.frame
units_var_table <- data.frame(variable = names(units_var), unit = units_var) 
```


## Convert to numeric
```{r}
for (i in var_num) data_keep[[i]] <- as.numeric(data_keep[[i]])
```


## Split the column 'Name' into several columns
```{r}
# these lines extract the artefact ID out of the path name
stud_name <- gsub(".* --- ", "", data_keep[["Name"]])
split_name <- do.call(rbind, strsplit(stud_name, "_"))

# splits the ID in the separat information 
data_final <- data.frame(split_name[,1], split_name[,2], split_name[,3], split_name[,4], split_name[,5], split_name[,6], data_keep[-3], stringsAsFactors = FALSE)
colnames(data_final)[1:9] <- c("Sample.ID", "Microscope", "Objective", "PolishType", "Surface", "Topo", "Acquisition.Date", "Analysis.Date", "Analysis.Time")

```

## Check the result
```{r}
str(data_final)
head(data_final)
```


---

# Save data
## Format name of output file
```{r}
file_out <- "data"
```
The files will be saved as "`r paste0("~/", file_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(data = data_final, units = units_var_table), 
           file = paste0(dir_out, file_out, ".xlsx"))
```


## Save R object
```{r}
saveObject(data_final, file = paste0(dir_out, file_out, ".Rbin"))
```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

