---
title: "Plots"
author: "Joao Marreiros"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  pdf_document: default
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```

__Brief description of the script__

This R markdown document plots the data of the resulting CSV file from the computing ISO 25178-2 parameters in ConfoMap. These data is part of the manuscript: *Dubreuil et al. A ‘family of wear’: Exploring use-wear patterns on ad hoc smoothing, burnishing/polishing tool*

This R project and respective scripts follow the procedures described by Marwick et al. 2017. 

The authors would like to thank Ivan Calandra and Lisa Schunk for their help and contribution on several chunks of code included here in the script (pieces of code are also adapted from Calandra et al. 2019, Pedergnana et al. 2020a, 2020b).

To compile this markdown document do not delete or move files from their original folders.

For any questions, comments and inputs, please contact:

Joao Marreiros, joao.marreiros@rgzm.de


---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(wesanderson)

dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
```

# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)
```

---


# Prepare variables
## Define numeric variables
```{r}
num.var <- 20:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep = "")
```

---


# Plot each of the selected numeric variables 

## Plot just original samples/surfaces organized by parameter
```{r}

for (i in num.var) {
	p <- ggplot(data = imp_data, aes_string(x = "Sample.ID", y = names(imp_data)[i],
	                                        colour = "PolishType")) + 
	     geom_point(size = 1) +
	     geom_jitter(width = 0.25) +  
	     theme_classic() +
	     labs(colour = "PolishType") +
	     labs(x = "Sample", y = gsub("\\.", " ", names(imp_data)[i])) +
	     scale_colour_hue(h = c(25,225), limits = levels(imp_data[["Surface"]]))
  print(p)
  
  # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_",
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf", width = 26,
	       height = 21, units = "cm" )
}
```

## Scatterplots of selected variables combined with Sample.ID and Polish type
### Sa vs. Sq
```{r}
#plot
Sa_Sq <- ggplot(data = imp_data) +  
         geom_point(mapping = aes(x = Sa, y = Sq, colour = PolishType)) +
         theme_classic() +
         theme(text = element_text(size = 14)) +
         labs(colour = "Surface") +
         facet_wrap(~ Sample.ID) +
         scale_colour_hue(h = c(25, 230)) 
print(Sa_Sq)

# saving the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_scatterplot_Sa-Sq", ".pdf")
ggsave(filename = file_out, plot = Sa_Sq, path = dir_out, device = "pdf")
```

### Sq vs. Vmc
```{r}
Sq_Vmc <- ggplot(data = imp_data) +  
          geom_point(mapping = aes(x = Sq, y = Vmc, colour = PolishType)) +
          theme_classic() +
          theme(text = element_text(size = 14)) +
          labs(colour = "Surface") +
          facet_wrap(~ Sample.ID) +
          scale_colour_hue(h = c(25, 230))  
print(Sq_Vmc)

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_scatterplot_Sq-Vmc", ".pdf")
ggsave(filename = file_out, plot = Sq_Vmc, path = dir_out, device = "pdf")
```

### Mean depth of furrows vs. mean density of furrows
```{r}
furrows <- ggplot(data = imp_data) +  
           geom_point(mapping = aes(x = Mean.depth.of.furrows, y = Mean.density.of.furrows, colour = PolishType)) +
           theme_classic() +
           theme(text = element_text(size = 14)) +
           labs(colour = "Surface", x = "Mean depth of furrows", y = "Mean density of furrows") +
           facet_wrap(~ Sample.ID) +
           scale_colour_hue(h = c(25, 230))
print(furrows)

file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_scatterplot_furrows", ".pdf")
ggsave(filename = file_out, plot = furrows, path = dir_out, device = "pdf")
```

---

# sessionInfo() and RStudio version
```{r}
sessionInfo()
```


